---
title: "PS03"
author: "Fatima Fairfax"
date: "9/4/21"
output:
  html_document:
    df_print: paged
---

##PS03: Data Wrangling

To set up, I'm running the relevant packages:
```{r}
library(ggplot2)
library(dplyr)
```

Next I will load and take a look at the data we're using:
```{r}
data(txhousing)
glimpse(txhousing)
```


**Question 1**

The last city listed in the data set is Wichita Falls


**Question 2**

The variable 'listings' correspond to the total active listings


Now I'm extracting the variables sales and volume:
```{r}
txhousing |> 
  select(sales,volume)
```

Now I'm looking at the data set with all variables *except* the date:
```{r}
txhousing <- txhousing |> 
  select(-date)
```

Filter to pull out specific rows based on criteria in the columns. First looking at all houses in the year 2012 in Austin:
```{r}
austin_12 <- txhousing |> 
  filter(year == 2012, city == "Austin")
```


Now all houses in Austin before the year 2004:
```{r}
txhousing |> 
  filter(year < 2004, city == "Austin")
```

Now houses in Austin or Abilene:
```{r}
aust_ab <- txhousing |> 
  filter(city == "Austin" | city == "Abilene")
```


Now I'll mutate to add a new variable that indicates volume in units of 100k:
```{r}
txhousing <- txhousing |> 
  mutate(vol_100k = volume/100000)
```

Now I'll provide summary stats for the Austin 2012 data to see the mean of sales:
```{r}
austin_12 |> summarize(x_bar_sales = mean(sales))
```

Now to look at a bunch of summaries at once:
```{r}
austin_12 |> summarize(x_bar_sales = mean(sales),
                       sd_sales = sd(sales), 
                        min_vol = min(volume), 
                        max_vol = max(volume), 
                        mdn_list = median(listings), 
                        iqr_list = IQR(listings),
                        sample_size = n())
```

Now I can arrange the data by the volume column:
```{r}
austin_12 |> 
  arrange(desc(volume))
```

Finally, I can use group_by to look at data across levels of a variable. In this case, looking at the average number of monthly sales in Austin and Abilene across all years:
```{r}
aust_ab |> group_by(city) |> 
  summarize(x_bar_sales = mean(sales))
```

I can also group_by multiple variables. This will show the average sales for each city for each month:
```{r}
aust_ab |> group_by(city, month) |> 
  summarize(x_bar_sales = mean(sales))
```


**Question 3**

I'll remove the inventory variable and overwrite the txhousing dataframe:
```{r}
txhousing <- txhousing |> select(-inventory)
```


**Question 4**

Now I'll make a subset of the data for Dallas in 2012 and 2013:
```{r}
dallas_sub <- txhousing |> filter(city == "Dallas", year == 2012 | year == 2013)
dallas_sub
```


**Question 5**

Now adding a new column to dallas_sub that calculated the percentage of listing sold:
```{r}
dallas_sub <- dallas_sub |>
  mutate(prct_sold = (sales/listings * 100))
dallas_sub
```

**Question 6**

Now calculating the average percentage of listing sold in each month of the year:
```{r}
dallas_summary <- dallas_sub |>
  group_by(month) |> 
  summarize(avg_prct_sold = mean(prct_sold))
dallas_summary
```


**Question 7**

Arranging the dallas_summary in desc order based on avg percent to show which had the largest. August (8) had the highest.
```{r}
dallas_summary |> arrange(desc(avg_prct_sold))
```

**Question 8**

```{r}
txhousing |>
  filter(year == 2012 | year == 2013, city == "Dallas") |> 
  mutate(prct_sold = sales/listings*100) |> 
  group_by(month) |> 
  summarize(mean_prct_sold = mean(prct_sold)) |> 
  arrange(desc(mean_prct_sold))
```

The above code did all of the operations I just ran separately above. Specifically, it took the original data frame and (1) filtered it to only include listings in Dallas in 2012 or 2013, (2) added a new column that was percentage of listing sold, (3) grouped the data by month, (4) created a summary value of the average percent sold and, (5) arranged the resulting data in descending order of percent sold.


**Question 9**

In Jan 2015, the city with the fewest houses listed for sale was: San Marcos
```{r}
txhousing |>
  filter(year == 2015, month == 1) |>
  arrange(listings)
```

**Question 10**

In 2012, the month where the most houses were sold was: August (8)
```{r}
txhousing |>
  filter(year == 2012) |> 
  group_by(month) |> 
  summarize(month_sales = sum(sales)) |> 
  arrange(desc(month_sales))
```

**Question 11**

Now generate a single table showing number of houses sold in Austin in 2000 and 2001 and the total of houses sold in Dallas in 2000 and 2001.

To do this, first I will select the relevant variables and filter for the cities and years I need to analyze: 
```{r}
aus_dall_0001 <- txhousing |>
  select(city, year,sales) |> 
  filter(city == "Austin" | city == "Dallas", year == 2000 | year == 2001)
aus_dall_0001
  
```

Then I will group_by city and sum up the sales:
```{r}
aus_dall_0001 <- aus_dall_0001 |> 
  group_by(city, year) |> 
  summarize(years_summed = sum(sales))
aus_dall_0001
```

Now I'll try grouping by city to combine the years:
```{r}
aus_dall_0001 |> 
  group_by(city) |> 
  summarize(total_homes_sold = sum(years_summed))
```

This makes me think I could have just grouped by city above to get to this table. since I already selected out for relevant cities and years. I'll try it that way below to see if that's right:

```{r}
aus_dall_v2 <- txhousing |>
  select(city, year,sales) |> 
  filter(city == "Austin" | city == "Dallas", year == 2000 | year == 2001) |> 
  group_by(city) |> 
  summarize(total_homes_sold = sum(sales))
aus_dall_v2
```

Yay! That worked. (Assuming these values are correct)





